---
title: "phenotype_data_lmer"
author: "Ruijuan Li"
date: "6/18/2018"
output: 
  html_document: 
    keep_md: yes
---

### Goal of this script:   
check whether "leave #", "internode #", "plant height (sum of internode length)", "leafl average", "leafw average", and "petiole length average" are significantly different between trt and gt 

### load libs 
```{r}
library(lmerTest) #loads lmerTest and lme4
library(tidyverse)
library(brms)
library(cowplot)
source("script/helper.R")  
```

### maximum likelihood method, lmer 

several Q: 

Q: why I decided only to show height, leaf length, and leaft width? A: Because they are the only shared phenotype measures between GH and GC 

```{r}
# 1) import data and reformat data 
setwd("~/Desktop/rapa_paper/Li_Rapa_Transcriptome_2018/")
GC.pheno <- read.csv("input/shade_avoidance_expt_in_GC.csv",na.string=c("NA","N/A")) # replace all NA w/ N/A (the standard missing data format in R)
GH.pheno <- read.csv("input/shade_avoidance_expt_in_GH.csv", na.string=c("NA","N/A")) 

head(GC.pheno)
head(GH.pheno)

summary(GC.pheno)
summary(GH.pheno)

names(GC.pheno)[3] <- "gt"
GC.pheno$rep <- as.factor(str_sub(GC.pheno$barcode,-2,-2))  # 8 reps 
GC.pheno$height <- GC.pheno$No..of.internodes * GC.pheno$intleng 
head(GC.pheno)
GC.pheno.data <- GC.pheno[, c("gt", "trt", "leaves", "No..of.internodes", "leafl", "leafw", "petleng", "rep", "height")]
colnames(GC.pheno.data) <- c("gt", "trt", "leaf_No", "internode_No", "leafl_ave", "leafw_ave", "petleng_ave", "rep", "height") 

GH.pheno$rep <- as.factor(GH.pheno$rep)
GH.pheno$petleng_ave <- round(rowMeans(subset(GH.pheno, select=c("pet1", "pet2", "pet3")), na.rm = T), digits = 1)
GH.pheno.data <- GH.pheno[, c("trt", "rep", "intleng", "leafleng", "leafwid", "leafnum", "internum", "species", "petleng_ave")]
colnames(GH.pheno.data) <- c("trt", "rep", "height", "leafl_ave", "leafw_ave", "leaf_No", "internode_No", "gt", "petleng_ave")
head(GH.pheno.data)

# 2) visualize outcome distribution 
pheno.plot(GC.pheno.data, GC.pheno.data$height) 
pheno.plot(GH.pheno.data, GH.pheno.data$height)

# 3) mixed-effect linear model selection (rep as random effect), using all original data scale
# GC
# R500_SUN as ref 
# set reference level 
GC.pheno.data$trt <- relevel(GC.pheno.data$trt, ref = "SUN")
GC.pheno.data$gt <- relevel(GC.pheno.data$gt, ref = "R500")

lmer.leafl_ave.GC <- lmer(leafl_ave ~ gt * trt + (1|rep), data = GC.pheno.data)
summary.leafl_ave.GC <- summary(lmer.leafl_ave.GC)$coefficient

lmer.leafw_ave.GC <- lmer(leafw_ave ~ gt * trt + (1|rep), data = GC.pheno.data)
summary.leafw_ave.GC <- summary(lmer.leafw_ave.GC)$coefficient

lmer.height.GC <- lmer(height ~ gt * trt + (1|rep), data = GC.pheno.data)
summary.height.GC <- summary(lmer.height.GC)$coefficient

# IMB211_SUN as ref
# set reference level 
GC.pheno.data$trt <- relevel(GC.pheno.data$trt, ref = "SUN")
GC.pheno.data$gt <- relevel(GC.pheno.data$gt, ref = "IMB211")

lmer.leafl_ave.GC.2 <- lmer(leafl_ave ~ gt * trt + (1|rep), data = GC.pheno.data)
summary.leafl_ave.GC.2 <- summary(lmer.leafl_ave.GC.2)$coefficient

lmer.leafw_ave.GC.2 <- lmer(leafw_ave ~ gt * trt + (1|rep), data = GC.pheno.data)
summary.leafw_ave.GC.2 <- summary(lmer.leafw_ave.GC.2)$coefficient

lmer.height.GC.2 <- lmer(height ~ gt * trt + (1|rep), data = GC.pheno.data)
summary.height.GC.2 <- summary(lmer.height.GC.2)$coefficient

# GH 
# R500_SUN as ref level 
GH.pheno.data$trt <- relevel(GH.pheno.data$trt, ref = "ND")
GH.pheno.data$gt <- relevel(GH.pheno.data$gt, ref = "R500")

lmer.height.GH <- lmer(height ~ gt * trt + (1|rep), data = GH.pheno.data)
summary.height.GH <- summary(lmer.height.GH)$coefficient

lmer.leafl_ave.GH <- lmer(leafl_ave ~ gt * trt + (1|rep), data = GH.pheno.data)
summary.leafl_ave.GH <- summary(lmer.leafl_ave.GH)$coefficient

lmer.leafw_ave.GH <- lmer(leafw_ave ~ gt * trt + (1|rep), data = GH.pheno.data)
summary.leafw_ave.GH <- summary(lmer.leafw_ave.GH)$coefficient

# IMB211_SUN as ref level 
GH.pheno.data$trt <- relevel(GH.pheno.data$trt, ref = "ND")
GH.pheno.data$gt <- relevel(GH.pheno.data$gt, ref = "IMB211")

lmer.height.GH.2 <- lmer(height ~ gt * trt + (1|rep), data = GH.pheno.data)
summary.height.GH.2 <- summary(lmer.height.GH.2)$coefficient

lmer.leafl_ave.GH.2 <- lmer(leafl_ave ~ gt * trt + (1|rep), data = GH.pheno.data)
summary.leafl_ave.GH.2 <- summary(lmer.leafl_ave.GH.2)$coefficient

lmer.leafw_ave.GH.2 <- lmer(leafw_ave ~ gt * trt + (1|rep), data = GH.pheno.data)
summary.leafw_ave.GH.2 <- summary(lmer.leafw_ave.GH.2)$coefficient

# 4) display lmer result 
plot.leafl_ave.GC <- pheno.plot.result.GC(summary.leafl_ave.GC, summary.leafl_ave.GC.2)
plot.leafw_ave.GC <- pheno.plot.result.GC(summary.leafw_ave.GC, summary.leafw_ave.GC.2)
plot.height.GC <- pheno.plot.result.GC(summary.height.GC, summary.height.GC.2)

plot.height.GH <- pheno.plot.result.GH(summary.height.GH, summary.height.GH.2)
plot.leafl_ave.GH <- pheno.plot.result.GH(summary.leafl_ave.GH, summary.leafl_ave.GH.2)
plot.leafw_ave.GH <- pheno.plot.result.GH(summary.leafw_ave.GH, summary.leafw_ave.GH.2)

plot.all.paper <-plot_grid(
  plot.height.GC+theme(legend.position="none",axis.text.x=element_blank())+labs(title="plant height", y="Growth Chamber\n mm"),
  plot.leafl_ave.GC+theme(legend.position="none",axis.text.x=element_blank())+labs(title="leaf length", y="mm"),
  plot.leafw_ave.GC+theme(legend.position="none",axis.text.x=element_blank())+labs(title="leaf width", y="mm", fill=""),
  plot.height.GH+theme(legend.position=c("none"),axis.text.x=element_blank())+labs(title=" ", y="Greenhouse\n mm"),
  plot.leafl_ave.GH+theme(legend.position="none",axis.text.x=element_blank())+labs(title=" ", y="mm"),
  plot.leafw_ave.GH+theme(legend.position="none",axis.text.x=element_blank())+labs(title=" ", y="mm", x="", fill=""),
  ncol=3, nrow = 2,labels=c("","","",""))  

legend <- get_legend(plot.height.GC)
plot.all.paper.final <- plot_grid(plot.all.paper, legend, rel_widths = c(6, 1))

# save_plot("output/figure1.png", plot.all.paper.final, ncol = 3, nrow = 2,base_aspect_ratio = 0.8)  
```

### Julin asked: 

Is the greater response of R500 was true both on an abosolute scale and relative to height under high R/FR? log2 transform the data to see the result, look at percentage change, this is based on log(yt)−log(yt−1)=log(yt/yt−1), and yt/yt−1 is 1 plus the percentage change, check this for plant height only  

```{r}
pheno.plot.log2(GC.pheno.data, GC.pheno.data$height)

# log2 transform GC. height 
GC.pheno.data$height_log2 <- log2(GC.pheno.data$height)

# fit model 
# R500_SUN as ref 
# set reference level 
GC.pheno.data$trt <- relevel(GC.pheno.data$trt, ref = "SUN")
GC.pheno.data$gt <- relevel(GC.pheno.data$gt, ref = "R500")
lmer.height.GC <- lmer(height_log2 ~ gt * trt + (1|rep), data = GC.pheno.data)
summary.height.GC <- summary(lmer.height.GC)$coefficient
summary(lmer.height.GC)
# Fixed effects:
#                   Estimate Std. Error      df t value Pr(>|t|)    
# (Intercept)         6.7081     0.1739 27.0000  38.575  < 2e-16 ***
# gtIMB211           -1.4618     0.2459 27.0000  -5.944 2.45e-06 ***
# trtSHADE            0.9951     0.2459 27.0000   3.909 0.000563 ***
# gtIMB211:trtSHADE  -0.3777     0.3540 27.0000  -1.067 0.295350  

GC.pheno.data$gt <- relevel(GC.pheno.data$gt, ref = "IMB211")
lmer.height.GC.2 <- lmer(height_log2 ~ gt * trt + (1|rep), data = GC.pheno.data)
summary.height.GC.2 <- summary(lmer.height.GC.2)$coefficient # shade becomes significant... although less 
summary(lmer.height.GC.2)
# Fixed effects: 
#                 Estimate Std. Error      df t value Pr(>|t|)    
# (Intercept)       5.2463     0.1739 27.0000  30.169  < 2e-16 ***
# gtR500            1.4618     0.2459 27.0000   5.944 2.45e-06 ***
# trtSHADE          0.6174     0.2546 27.0000   2.425   0.0223 *  
# gtR500:trtSHADE   0.3777     0.3540 27.0000   1.067   0.2954    

# model result 
plot.height.GC <- pheno.plot.result.GC(summary.height.GC, summary.height.GC.2)
plot.height.GC # not significant anymore... 
save_plot("~/Desktop/test.png", plot.height.GC, base_aspect_ratio = 1.2)   
```

### bayesain method, brms  
```{r}

```

